---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The following is based off the public data set provided by the Seattle Public Library. Due to the sheer size of the data, this specific report uses the data filtered to items that have been checked out more than 10 times from 2017-2023.

The first chart is inspired by J.K. Rowling's popular series Harry Potter. This series has become well known to all since beginning as a book series in 1998 and being adapted in a film series in 2001. Since some time has passed since then, how relevant is this series still? 

The second chart aims to understand just exactly how technology and world events have influenced content consumption. How popular have different media forms been throughout these past five years?

The last chart demonstrates what books were popular in which years and just how popular they were simply by comparing the checkout amount to each year. Can we connect the data to known world events that triggered their popularity?

### Summary Information

Many people underestimate the resources that the library provides. Some say that with the direction the technology is going, the internet may be all we need and libraries will no longer serve a purpose. To demonstrate that this is not the case at all, here are the total amount of digital and physical items that were checked out from in 2017 and 2022:

- 2017
  - Digital: 756,756
  - Physical: 2,763,572
- 2022
  - Digital: 1,962,294
  - Physical: 1,094,618

While physical items are being checked out by almost half the amount the once were, digital checkouts have increased by almost 2.6 times. The SPL has done an excellent job of keeping up with technological changes and continuing to provide accessible resources. 



### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning = FALSE, code = readLines("chart1.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, code = readLines("chart2.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, fig.width=10, fig.height=10, echo = FALSE, code = readLines("chart3.R")}
```


